Improving real-time trend estimates using local parametrisation of polynomial regression filters

### 2024-04-25 paper/article_jos.tex
update article

This paper studies and compares real-time estimates of the trend-cycle
component with moving averages based on local polynomial regression. In
particular, this theory allows us to reproduce Henderson's symmetric and
Musgrave's asymmetric filters. This paper describes how they can be
extended to include a timeliness criterion to minimise the phase shift.
While asymmetric filters are generally parametrised globally, which can
be suboptimal around turning points, this paper proposes a procedure for
parametrising them locally. An empirical comparison, based on simulated
and real data, shows that modelling polynomial trends that are too
complex introduces more revisions without reducing the phase shift, and
that local parametrisation reduces the delay in detecting turning points
and reduces revisions. The paper also shows how these results can be
easily reproduced using the R package \texttt{rjd3filters} dedicated to
the manipulation of moving averages.

@@

Using local polynomial regression, \textcite{proietti2008} developed a
general class of symmetric and asymmetric moving averages. In
particular, it becomes possible to reproduce Henderson's symmetric
filter and Musgrave's asymmetric filters. However, these methods have
two drawbacks. Firstly, the phase shift (i.e.~the delay in detecting
turning points) is not directly controlled. Secondly, asymmetric filters
depend on a parameter that is generally estimated over the whole series
using assumptions that may be wrong locally. The aim of this study is to
propose extensions to the \textcite{proietti2008} class of filters in
order to take these shortcomings into account. The extensions have been
implemented in the statistical software R \autocite{logicielr}
\texttt{rjd3filters} package (available at
\url{https://github.com/rjdemetra/rjd3filters}) and can therefore be
easily used.

In section~\ref{sec-propMM}, we describe the general properties of
moving averages and the associated quality criteria. This allows us to
understand the foundations behind the construction of moving averages
from local polynomial regressions, as well as those behind the two
extensions proposed in this article (section~\ref{sec-nonparamreg}). In
a final section (section~\ref{sec-comparison}), all these methods are
compared empirically on simulated and real series.

\section{Some properties of moving averages}\label{sec-propMM}

@@

Let \(p\) and \(f\) two integers, a moving average
\(M_{\boldsymbol\theta}\) or \(M\) is defined by a set of coefficients
\(\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}\)
such as for all time series \(X_t\): \[
M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
\]

@@

To interpret the notions of gain and phase shift, it is useful to
illustrate the effects of moving averages on harmonic series
\(X_t=\e^{-i\omega t}\) with \(\omega\in[0,\pi]\). The moving average
\(M_{\boldsymbol\theta}\) transforms \(X_t\) into: \[
Y_t = M_{\boldsymbol\theta}X_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot X_t.
\] The function
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}\)
is called the \emph{transfer function} or \emph{frequency response
function}. The frequency response function can alternatively be defined
as
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}\)
or
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}\).
It can be rewritten as: \[
\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
\] where
\(\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert\)
is the \emph{gain} or \emph{amplitude} function and
\(\varphi_{\boldsymbol\theta}(\omega)\) is the \emph{phase shift} or
\emph{time shift} function. We this have: \begin{align*}
\rho_{\boldsymbol\theta}(\omega) &=\sqrt{\left(\sum_{k=-p}^{+f}\theta_k\cos(\omega k)\right)^2+
\left(\sum_{k=-p}^{+f}\theta_k\sin(\omega k)\right)^2} \\
\tan \varphi_{\boldsymbol\theta}(\omega) &=- \frac{\sum_{k=-p}^{+f}\theta_k\sin(\omega k)}{\sum_{k=-p}^{+f}\theta_k\cos(\omega k)} \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) \ne 0 \\
&= 0 \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) = 0.
\end{align*} The phase shift is sometimes represented as
\(\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}\)
to measure it in number of periods. For all symmetric moving averages we
have \(\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})\).

To sum up, applying a moving average to a harmonic times series
(\(X_t=\e^{-i\omega t}\)) affects it in in two different ways:

@@

Fourier decomposition allows us to analyze any time series as a sum of
harmonic series, and each component (trend, cycle, seasonal, irregular)
is associated with a set of frequencies. Noting \(\omega = 2\pi/s\), the
harmonic series of frequency \(\omega\) represents a series that repeats
itself every \(s\) periods. For example, for a monthly series (\(s=12\)
observations per year), seasonal movements are those that repeat every
year: they are therefore associated with frequencies \(2\pi/12\) (annual
periodicity), \(2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5\). In this
paper, we consider that the trend-cycle is associated with frequencies
in the interval \([0, 2\pi/12]\), i.e.~movements recurring at least
every 12 months. Even if different frequencies are sometimes used
(e.g.~\([0, 2\pi/36]\) to consider only cycles of at least 36 months),
this has no impact on the results of the study. The other frequencies,
\([2\pi/12, \pi]\) are associated with the irregular component
(undesirable oscillations).

figure~\ref{fig-graphsmusgrave} shows the gain and phase shift function
for the asymmetric Musgrave filter (see section~\ref{sec-lppasymf})
often used for real-time trend-cycle estimation (i.e.~when no point in
the future is known). The gain function is greater than 1 on the
frequencies associated with the trend-cycle (\([0, 2\pi/12]\)): this
means that the trend-cycle is well preserved and that short cycles of 1
to 2 years (\([2\pi/24,2\pi/12]\)) are even amplified. At other
frequencies, the gain function is less than 1, but always positive: this
means that the series smoothed by this moving average will always
contain noise, even if it is attenuated. Analysis of the phase shift
shows that the shorter the cycles, the greater the phase shift: this
means that on series smoothed by this moving average, turning points
could be detected at the wrong date.

\caption{\label{fig-graphsmusgrave}Coefficients, gain and phase shift
function for the Musgrave filter in real time with \(I/C=3.5\).}

@@

@@

In our case, we assume that our initial series \(y_t\) is seasonally
adjusted and can be written as the sum of a trend-cycle, \(TC_t\), and
an irregular component, \(I_t\): \[
y_t=TC_t+I_t.
\] The aim will be to build moving averages that best preserve the
trend-cycle (\(M_{\boldsymbol\theta} (TC_t)\simeq TC_t\)) and minimise
noise (\(M_{\boldsymbol\theta} (I_t)\simeq 0\)).

\subsubsection{Trend preservation}\label{trend-preservation}

Trend-cycles are generally modelled by local polynomial trends (see
section~\ref{sec-nonparamreg}). In order to best preserve the
trend-cycles, we try to design moving averages that preserve the
polynomial trends. A moving average \(M_{\boldsymbol\theta}\) preserves
a function of time \(f(t)\) if
\(\forall t:M_{\boldsymbol\theta} f(t)=f(t)\).

@@

\subsubsection{Noise reduction}\label{noise-reduction}

All time series are affected by noise, which can blur the extraction of
trends and cycles. This is why we try to reduce this noise (by reducing
its variance) while preserving the relevant trends. The sum of the
squares of the coefficients \(\sum_{k=-p}^{+f}\theta_k^2\) is the
\emph{reduction of variance} ratio.

Indeed, let \(\{\varepsilon_t\}\) be a sequence of independent random
variables with \(\E{\varepsilon_t}=0\), \(\V{\varepsilon_t}=\sigma^2\):
\[
\V{M_{\boldsymbol\theta}\varepsilon_t}=\V{\sum_{k=-p}^{+f} \theta_k \varepsilon_{t+k}}
= \sum_{k=-p}^{+f} \theta_k^2 \V{\varepsilon_{t+k}}=
\sigma^2\sum_{k=-p}^{+f} \theta_k^2.
\]

@@

\subsubsection{Asymmetric moving averages and
forecasts}\label{sec-mmetprev}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Use asymmetric moving averages to take account of the lack of
  available data;
\item
  Apply symmetric filters to extended forecast series. This method seems
  to date back to \textcite{deforest1877adjustment}, which also suggests
  modelling a polynomial trend of degree three or less at the end of the
  period. A similar approach is used in the X-13ARIMA-SEATS seasonal
  adjustment method: the series is extend by one year (by default) with
  an ARIMA model to minimise the revisions associated to asymmetric
  filters
\end{enumerate}

@@

Various estimation methods can be used to derive symmetric and
asymmetric moving averages.\\
\textcite{GrayThomson1996} propose a complete statistical framework that
makes it possible, in particular, to model the error in approximating
the trend by local polynomials. However, as the specification of this
error is generally complex, simpler models may be preferred, such as
that of \textcite{proietti2008}.\\
Finally, \textcite{dagumbianconcini2008} propose a similar modelling of
the trend-cycle but using the theory of Hilbert spaces with reproducing
kernels for estimation, which has the particular advantage of
facilitating the calculation of different moving averages at different
time frequencies, this method can be used in R with the function
\texttt{rjd3filters::rkhs\_filter()}.

\subsection{Symmetric moving averages and local polynomial
regression}\label{symmetric-moving-averages-and-local-polynomial-regression}

@@

As mentioned in section~\ref{sec-mmetprev}, for real-time estimation,
several approaches can be used:

@@

Rewriting equation~\ref{eq-lpp}:
\begin{equation}\phantomsection\label{eq-lpgeneralmodel}{
\boldsymbol y=\begin{pmatrix}\boldsymbol U &\boldsymbol Z\end{pmatrix}
\begin{pmatrix}\boldsymbol \gamma \\ \boldsymbol \delta\end{pmatrix}+\boldsymbol \varepsilon,\quad
\boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
}\end{equation} where \([\boldsymbol U,\boldsymbol Z]\) is of full rank
and forms a subset of the columns of \(X.\) Defining the variance error
as \(\boldsymbol D = \sigma^2\boldsymbol K^{-1}\) (to directly take into
account the weights associated to the observations) and with
\([\boldsymbol U,\boldsymbol Z] = \boldsymbol X\) we this find the model
associated to symmetric filters. The objective is to find an asymmetric
filter \(\boldsymbol v\) which uses \(q\) points in the future and
minimises the mean square error of revision (to the symmetric filter
\(\boldsymbol \theta\)) under certain (polynomial) constraints. The
matrices \(\boldsymbol y\), \(\boldsymbol U\), \(\boldsymbol Z\) and
\(\boldsymbol \theta\) are partitioned as follows: \[
\boldsymbol y = \begin{pmatrix} \boldsymbol y_p \\ \boldsymbol y_f \end{pmatrix}, \quad
\boldsymbol U = \begin{pmatrix} \boldsymbol U_p \\ \boldsymbol U_f \end{pmatrix}, \quad  
\boldsymbol Z = \begin{pmatrix} \boldsymbol Z_p \\ \boldsymbol Z_f \end{pmatrix}, \quad 
\boldsymbol \theta = \begin{pmatrix} \boldsymbol \theta_p \\ \boldsymbol \theta_f \end{pmatrix},
\] where \(\boldsymbol y_p\) is a matrix \((h+q+1)\times 1\) which
denotes the set of available observations (from \(y_{t-h}\) to
\(y_{t+q}\)) and \(\boldsymbol U\), \(\boldsymbol Z\) and
\(\boldsymbol \theta\) are partitioned in the same way. The constraints
are defined as
\(\transp{\boldsymbol U_p}\boldsymbol v=\transp{\boldsymbol U}\boldsymbol \theta\):
they represents the polynomial preservation constraints we want to
impose on the asymmetric filter. \textcite{proietti2008} show that the
problem is equivalent to finding \(v\) which minimises:
\begin{equation}\phantomsection\label{eq-lppasym}{
\varphi(\boldsymbol v)=
\underbrace{
  \underbrace{\transp{(\boldsymbol v-\boldsymbol \theta_{p})}\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
  \transp{\boldsymbol \theta_{f}}\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
  \underbrace{[\transp{\boldsymbol \delta}(\transp{\boldsymbol Z_{p}}\boldsymbol v-\transp{\boldsymbol Z}\boldsymbol \theta)]^{2}}_{bias^2}
}_\text{mean square revision error}+
\underbrace{2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)}_{\text{constraints}}.
}\end{equation} where \(\boldsymbol l\) is a vector of Lagrange
multipliers.

When \(\boldsymbol U=\boldsymbol X\) (and then
\(\boldsymbol Z = \boldsymbol 0\)), the constraint is equivalent to
preserving polynomials of degree \(d\) and we find asymmetric direct
filters (DAF) when \(\boldsymbol D=\boldsymbol K^{-1}\).

@@

Appendix \ref{sec-an-cof} shows the coefficients, gain and phase shift
functions of the four asymmetric filters.

@@

In the empirical applications of section~\ref{sec-comparison}, the final
local parametrisation is obtained by estimating \(\delta\) using all the
data (i.e., using the symmetric filters shown in
figure~\ref{fig-mmpenteconcac}), while maintaining a real-time estimate
of \(\sigma^2\). figure~\ref{fig-mmpenteconcac-ex} shows an example of
the comparison of estimates of the ratio \(|\delta_1/\sigma|\) with a
global estimator (IC ratio) and two local parametrisations:

@@

The turning points are clearly detected by the local estimators, but
there is much more variance in the estimate of \(|\delta_1/\sigma|\) in
the real-time estimate. This will lead to another source of revisions in
the intermediate estimates.

@@

@@

The different methods are compared on simulated and real data. For all
series, a symmetric 13-term filter is used: we then assume than the
bandwidth is fixed, as in the seasonal adjustement method X-11. A
nearest neighbour bandwidth is also tested (13-term asymmetric filters:
real-time estimates uses observation between \(y_{t-12}\) and \(y_{t}\),
the second asymmetric filters uses obversations between \(y_{t-11}\) and
\(y_{t+1}\), etc.) but with no improvement in the results (see
supplemental material). More complexe methods to locally select
bandwidth, like the one proposed by \textcite{fan1992variable}, were not
tested. All the methods to create asymmetric filters are also compared
with estimates obtained by extending the series with an ARIMA model,
then applying a 13-term symmetric Henderson filter. The ARIMA model is
determined automatically, using no seasonal lag (the series being
seasonally adjusted) and no external variables (such as external
regressors for correction of atypical points). The performance of the
different methods is judged on criteria relating to revisions (between
two consecutive estimates and in relation to the final estimate) and the
number of periods required to detect turning points.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Computation of the phase shift in the detection of turning points. The
  definition of \textcite{Zellner1991} is used to determine the turning
  points:

>>>>

This paper studies and compares real-time estimates of the trend-cycle
component with moving averages based on local polynomial regression. In
particular, this theory allows us to reproduce Henderson's symmetric and
Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal
adjustment algorithm. This paper proposes two extensions of local
polynomial filters for real-time trend-cycle estimates (for the last
estimates): first including a timeliness criterion to minimise the phase
shift; second, with procedure for parametrising asymmetric filters
locally While they are generally parametrised globally, which can be
suboptimal around turning points. An empirical comparison, based on
simulated and real data, shows that modelling polynomial trends that are
too complex introduces more revisions without reducing the phase shift,
and that local parametrisation reduces the delay in detecting turning
points and reduces revisions. The paper also shows how these results can
be easily reproduced using the R package \texttt{rjd3filters} dedicated
to the manipulation of moving averages.

@@

Using local polynomial regression, \textcite{proietti2008} developed a
general class of symmetric and asymmetric moving averages. In
particular, it becomes possible to reproduce Henderson's symmetric
filter and Musgrave's asymmetric filters. The aim of this study is to
propose two extensions to the \textcite{proietti2008} class of filters
in order to two drawbacks of those methods. Firstly, adding, in the
optimisation process, a timeliness criterin to directly control the
phase shift (i.e.~the delay in detecting turning points). Secondly,
calibrating asymmetric filter locally, while they are generally
calibrated globally using thus assumptions that may be wrong locally.
Those extensions are implemented in the statistical software R
\autocite{logicielr} throw the \texttt{rjd3filters} package (available
at \url{https://github.com/rjdemetra/rjd3filters}) and can therefore be
easily used. All the code associated to this paper is available at
\url{https://github.com/XXXX} (anonymized linked before the manuscript
is accepted).

In section~\ref{sec-propMM}, we describe the general properties of
moving averages and the associated quality criteria. This allows us to
understand the foundations behind the construction of moving averages
from local polynomial regressions, as well as those behind the two
extensions proposed in this article (section~\ref{sec-nonparamreg}). In
section~\ref{sec-comparison}, all these methods are compared empirically
on simulated and real series.

\section{Local trend-cycle models and moving averages}\label{sec-propMM}

The basic assumption used in time series decomposition methods is that
the input observed time series, \(y_t\) can be decomposed into a signal
component \(\mu_t\) and a an erratic component \(\varepsilon_t\) (called
irregular): \[
f(y_t)=\mu_t+\varepsilon_t
\] where \(f\) denotes some appropriate transformation (usually
logarithmic or no transformation). To simplify further notations,
\(y_t\) will denote the transformed observed series. The noise component
\(\varepsilon_t\) is usually assumed to be a white noise. Assuming that
the input time series is seasonally adjusted (or without seasonality),
the signal \(\mu_t\) represents the trend and cycle components, here
estimated jointly and called trend-cycle \(TC_t\). This component can be
locally approximated by a polynomial degree (see
section~\ref{sec-nonparamreg}): \[
\mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
\] with \(\xi_t\) a stochastic process uncorrelated with
\(\varepsilon_t\). Even if some papers model \(\xi_t\) and
\(\varepsilon_t\) separately \autocite[see for
example][]{GrayThomson2002}, an usual assumption, adopted in this paper,
is to is to gather \(\xi_t\) and \(\varepsilon_t\). Thus, the
trend-cycle \(TC_t\) is considered as deterministic and modelled as a
polynomial trend of degree \(d\). The coefficients
\((\beta_0,\dots,\beta_k)\) can be estimated by weighted least squares.
The estimate \(\hat \beta_0\) provides the trend-cycle estimate
\(\hat TC_t\) and it can be shown that this equivalent to apply of
moving average (see section~\ref{sec-nonparamreg}). The goal of this
paper is to present two new methods to separate the trend-cycle
component from the noise component in real-time (i.e.~when no future
point is known), using moving averages build on local polynomial
smoothing techniques.

@@

Let \(p\) and \(f\) two integers, a moving average
\(M_{\boldsymbol\theta}\) or \(M\) is defined by a set of coefficients
\(\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}\)
such as for all time series \(y_t\): \[
M_{\boldsymbol\theta}(y_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
\]

@@

To interpret the notions of gain and phase shift, it is useful to
illustrate the effects of moving averages on harmonic series
\(y_t=\e^{-i\omega t}\) with \(\omega\in[0,\pi]\). The moving average
\(M_{\boldsymbol\theta}\) transforms \(y_t\) into: \[
Y_t = M_{\boldsymbol\theta}y_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot y_t.
\] The function
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}\)
is called the \emph{transfer function} or \emph{frequency response
function}. The frequency response function can alternatively be defined
as
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}\)
or
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}\).
It can be rewritten as: \[
\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
\] where
\(\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert\)
is the \emph{gain} or \emph{amplitude} function and
\(\varphi_{\boldsymbol\theta}(\omega)\) is the \emph{phase shift} or
\emph{time shift} function. We this have: \begin{align*}
\rho_{\boldsymbol\theta}(\omega) &=\sqrt{\left(\sum_{k=-p}^{+f}\theta_k\cos(\omega k)\right)^2+
\left(\sum_{k=-p}^{+f}\theta_k\sin(\omega k)\right)^2} \\
\tan \varphi_{\boldsymbol\theta}(\omega) &=- \frac{\sum_{k=-p}^{+f}\theta_k\sin(\omega k)}{\sum_{k=-p}^{+f}\theta_k\cos(\omega k)} \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) \ne 0 \\
&= 0 \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) = 0.
\end{align*} The phase shift is sometimes represented as
\(\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}\)
to measure it in number of periods. For all symmetric moving averages we
have \(\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})\).

To sum up, applying a moving average to a harmonic times series
(\(y_t=\e^{-i\omega t}\)) affects it in in two different ways:

@@

Fourier decomposition allows us to analyze any time series as a sum of
harmonic series, and each component (trend, cycle, seasonal, irregular)
is associated with a set of frequencies. Noting \(\omega = 2\pi/s\), the
harmonic series of frequency \(\omega\) represents a series that repeats
itself every \(s\) periods. For example, for a monthly series (\(s=12\)
observations per year), seasonal movements are those that repeat every
year: they are therefore associated with frequencies \(2\pi/12\) (annual
periodicity), \(2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5\). In this
paper, we consider that the trend-cycle is associated with frequencies
in the interval \([0, 2\pi/12]\), i.e.~movements recurring at least
every 12 months. Even if different frequencies are sometimes used
(e.g.~\([0, 2\pi/36]\) to consider only cycles of at least 36 months),
this has no impact on the results of the study. The other frequencies,
\([2\pi/12, \pi]\) are associated with the irregular component.

Figure~\ref{fig-graphsmusgrave} shows the gain and phase shift function
for the asymmetric Musgrave filter (see section~\ref{sec-lppasymf})
often used for real-time trend-cycle estimation (i.e.~when no point in
the future is known). The gain function is greater than 1 on the
frequencies associated with the trend-cycle (\([0, 2\pi/12]\)): this
means that the trend-cycle is well preserved and that short cycles of 1
to 2 years (\([2\pi/24,2\pi/12]\)) are even amplified. At other
frequencies, the gain function is less than 1, but always positive: this
means that the series smoothed by this moving average will always
contain noise, even if it is attenuated. Analysis of the phase shift
shows that the shorter the cycles, the greater the phase shift: this
means that on series smoothed by this moving average, turning points
could be detected at the wrong date.

\begin{figure}

\caption{\label{fig-graphsmusgrave}Coefficients, gain and phase shift
function for the Musgrave filter used in X-11 in real time (last
estimate) for monthly series when the bandwidth is 6.}

@@

Note: the Musgrave filter depends on a ratio \(I/C\) (see
section~\ref{sec-nonparamreg}). In X-11, for monthly series, when the
bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used,
what is used in most cases), the ratio is fixed to 3.5 and this value is
used in this figure.

@@

In this paper, we assume that our initial series \(y_t\) is seasonally
adjusted and can be written as the sum of a trend-cycle, \(TC_t\), and
an irregular (noise) component, \(\varepsilon_t\): \[
y_t=TC_t+\varepsilon_t.
\] The aim will be to build moving averages that best preserve the
trend-cycle (\(M_{\boldsymbol\theta} (TC_t)\simeq TC_t\)).

Trend-cycles are generally modelled by local polynomial trends (see
section~\ref{sec-nonparamreg}) and, in order to best preserve the
trend-cycles, we thus design moving averages that preserve the
polynomial trends. A moving average \(M_{\boldsymbol\theta}\) preserves
a function of time \(f(t)\) if
\(\forall t:M_{\boldsymbol\theta} f(t)=f(t)\).

@@

@@

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Use asymmetric moving averages to take account of the lack of
  available data;
\item
  Apply symmetric filters to extended forecast series. This method seems
  to date back to \textcite{deforest1877adjustment}, which also suggests
  modelling a polynomial trend of degree three or less at the end of the
  period. A similar approach is used in the X-13ARIMA-SEATS seasonal
  adjustment method: the series is extend by one year (by default) with
  an ARIMA model to minimise the revisions associated to asymmetric
  filters. Since the final filter used for the seasonal adjustment
  extraction is much longer than one year (in most of the cases, for
  monthly series, the bandwidth is 78), the input series is just
  partially extended (and asymmetric filters are still used) to prevent
  from instability over long periods (e.g., outliers or model shifts).
\end{enumerate}

@@

Various estimation methods can be used to derive symmetric and
asymmetric moving averages.\\
\textcite{GrayThomson1996} propose a complete statistical framework that
makes it possible, in particular, to model the error in approximating
the trend by local polynomials. However, as the specification of this
error is generally complex, simpler models may be preferred, such as
that of \textcite{proietti2008}.\\
Finally, \textcite{dagumbianconcini2008} propose a similar modelling of
the trend-cycle but using the theory of Hilbert spaces with reproducing
kernels for estimation, which has the particular advantage of
facilitating the calculation of different moving averages at different
time frequencies. See \textcite{inseeDTM202401} for a more detailed
comparison of the different recent methods for trend-cycle extraction
included in the \texttt{rjd3filters} package

In this paper we will focus on \textcite{proietti2008} approach.
Section~\ref{sec-sym-lp} and section~\ref{sec-lppasymf} describes their
approach to build symmetric moving averages (used for the final
estimates of the trend-cycle component) and asymmetric moving averages
(used for real-time estimates). This paper proposed two extensions of
their approach to build asymetric moving averages: a first one adding a
criteria to control the phase shift (section~\ref{sec-localic}), not
used in the simulation (section~\ref{sec-comparison}) but which could be
used in future studied; a second one using a local parameterisation of a
parameter usually parameterised globally
(section~\ref{sec-lptimeliness}), which is used in the simulation.

\subsection{Symmetric moving averages and local polynomial
regression}\label{sec-sym-lp}

@@

As mentioned in section~\ref{sec-mmasym}, for real-time estimation,
several approaches can be used:

@@

Rewriting equation~\ref{eq-lpp}:
\begin{equation}\phantomsection\label{eq-lpgeneralmodel}{
\boldsymbol y=\begin{pmatrix}\boldsymbol U &\boldsymbol Z\end{pmatrix}
\begin{pmatrix}\boldsymbol \gamma \\ \boldsymbol \delta\end{pmatrix}+\boldsymbol \varepsilon,\quad
\boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
}\end{equation} where \([\boldsymbol U,\boldsymbol Z]\) is of full rank
and forms a subset of the columns of \(X.\) Defining the variance error
as \(\boldsymbol D = \sigma^2\boldsymbol K^{-1}\) (to directly take into
account the weights associated to the observations) and with
\([\boldsymbol U,\boldsymbol Z] = \boldsymbol X\) we thus find the model
associated to symmetric filters. The objective is to find an asymmetric
filter \(\boldsymbol v\) which uses \(q\) points in the future and
minimises the mean square error of revision (to the symmetric filter
\(\boldsymbol \theta\)) under certain (polynomial) constraints. The
matrices \(\boldsymbol y\), \(\boldsymbol U\), \(\boldsymbol Z\) and
\(\boldsymbol \theta\) are partitioned as follows: \[
\boldsymbol y = \begin{pmatrix} \boldsymbol y_p \\ \boldsymbol y_f \end{pmatrix}, \quad
\boldsymbol U = \begin{pmatrix} \boldsymbol U_p \\ \boldsymbol U_f \end{pmatrix}, \quad  
\boldsymbol Z = \begin{pmatrix} \boldsymbol Z_p \\ \boldsymbol Z_f \end{pmatrix}, \quad 
\boldsymbol \theta = \begin{pmatrix} \boldsymbol \theta_p \\ \boldsymbol \theta_f \end{pmatrix},
\] where \(\boldsymbol y_p\) is a matrix \((h+q+1)\times 1\) which
denotes the set of available observations (from \(y_{t-h}\) to
\(y_{t+q}\)) and \(\boldsymbol U\), \(\boldsymbol Z\) and
\(\boldsymbol \theta\) are partitioned in the same way. The constraints
are defined as
\(\transp{\boldsymbol U_p}\boldsymbol v=\transp{\boldsymbol U}\boldsymbol \theta\):
they represents the polynomial preservation constraints we want to
impose on the asymmetric filter. \textcite{proietti2008} show that the
problem is equivalent to finding \(v\) which minimises:
\begin{equation}\phantomsection\label{eq-lppasym}{
\varphi(\boldsymbol v)=
\underbrace{
  \underbrace{\transp{(\boldsymbol v-\boldsymbol \theta_{p})}\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
  \transp{\boldsymbol \theta_{f}}\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
  \underbrace{[\transp{\boldsymbol \delta}(\transp{\boldsymbol Z_{p}}\boldsymbol v-\transp{\boldsymbol Z}\boldsymbol \theta)]^{2}}_{bias^2}
}_\text{mean square revision error}+
\underbrace{2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)}_{\text{constraints}}.
}\end{equation} where \(\boldsymbol l\) is a vector of Lagrange
multipliers.

When \(\boldsymbol U=\boldsymbol X\) (and then
\(\boldsymbol Z = \boldsymbol 0\)), the constraint is equivalent to
preserving polynomials of degree \(d\) and we find asymmetric direct
filters (DAF) when \(\boldsymbol D=\sigma^2\boldsymbol K^{-1}\).

@@

Supplemental materials shows the coefficients, gain and phase shift
functions of the four asymmetric filters.

@@

To distinguish the gain from the local parametrisation of the noise
biased associated to the real-time estimates of the slope and concavity,
in the empirical applications of section~\ref{sec-comparison} we also
compares the results to the final local parametrisation obtained by
estimating \(\delta\) using all the data (i.e., using the symmetric
filters shown in figure~\ref{fig-mmpenteconcac}), while maintaining a
real-time estimate of \(\sigma^2\). figure~\ref{fig-mmpenteconcac-ex}
shows an example of the comparison of estimates of the ratio
\(|\delta_1/\sigma|\) with a global estimator (IC ratio) and two local
parametrisations:

@@

The turning points are clearly detected by the local estimators
(changement in the slope), when \(|\delta_1/\sigma|\) is estimated in
the real-time estimate. However, the real-time local estimates of
\(|\delta_1/\sigma|\) are noisy: the varability of this estimates, used
to build asymmetric filters, will then lead to another source of
revisions in the intermediate estimates.

@@

There is no function in \texttt{rjd3filters} to directly compute those
moving averages but they can be easily computed using matrix computation
formula and applied to the series rjd3filters`.

@@

The different methods are compared on simulated and real data. For all
series, a symmetric 13-term filter is used: we then assume than the
bandwidth is fixed, as in the seasonal adjustement method X-11. A
nearest neighbour bandwidth is also tested (13-term asymmetric filters:
real-time estimates uses observation between \(y_{t-12}\) and \(y_{t}\),
the second asymmetric filters uses obversations between \(y_{t-11}\) and
\(y_{t+1}\), etc.) but with no improvement in the results (see
supplemental material). More complexe methods to locally select
bandwidth, like the one proposed by \textcite{fan1992variable}, were not
tested. All the methods to create asymmetric filters are also compared
with estimates obtained by extending the series with an ARIMA model,
then applying a 13-term symmetric Henderson filter. The ARIMA model is
determined automatically, using \textcite{autoarima} algorithm (function
\texttt{forecast::auto.arima()}) on the last 12 years, with no seasonal
lag (the series being seasonally adjusted) and no external variables
(such as external regressors for correction of atypical points).

The performance of the different methods is judged on criteria relating
to revisions (between two consecutive estimates and in relation to the
final estimate) and the number of periods required to detect turning
points.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Computation of the phase shift in the detection of turning points. In
  this paper, we focus on turning points associated to business cycle:
  it is defined as the succession of phases of economic recession and
  expansion, delimited by peaks (highest level of activity) and troughs
  (lowest level of activity), see for example \textcite{ferrara} for a
  description of the different economic cycles. In this paper, following
  \textcite{Zellner1991}, we used an accepted definition for
  turning-points:

  \begin{itemize}
  \tightlist
  \item
    An upturn occurs when the economy moves from a phase of recession to
    a phase of expansion. This is the case at date \(t\) when
    \(TC_{t-3}\geq TC_{t-2}\geq TC_{t-1}<TC_t\leq TC_{t+1}\).\\
  \item
    A downturn occurs when the economy moves from a phase of expansion
    to a phase of recession. This is the case at date \(t\) when
    \(TC_{t-3}\leq TC_{t-2}\leq TC_{t-1}>TC_t\geq TC_{t+1}\).
  \end{itemize}
